---
title: "Exercises"
author: "Andy Pham"
date: "April 21, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE, autodep=TRUE)
library(ggplot2)
```

##Exercise 1

```{r Exercise_1}
sum(2000:20000)
```
The sum is: 198011000

##Exercise 2
In the code snippet, a is assigned the value 5, b is assigned the vector from that holds the numbers 2 to 20 incremented by 1. a and b are then printed out.

##Exercise 3

```{r Exercise_3}
a <- 5
b <- 2:20
sum(a, b)
a + b
```
I get different results because sum adds all the numbers stored in both a and b while the "+" adds a to every element in b because b is a vector.

##Exercise 4

```{r Exercise_4}
sum(b[5:10])
```
The sum is 51.

##Exercise 5

```{r Exercise_5}
sum(b[c(3, 8, 10)])
```
The sum is 24.

##Exercise 6

```{r Exercise_6}
m <- matrix(data=1:25, ncol=5, byrow=T)
m
c(m[, 3], m[, 4], m[, 5])
```
When extracting from a 2D object, the first number represents the row and the second number represents the column. m[3,] will return all the values in row 3. You extract the columns of m together using c(m[, 3], m[, 4], m[, 5]).

##Exercise 7

```{r Exercise_7}
cbind(m, 101:105)
n <- 1:5
rbind(n, m)
```
The cbind command adds a column to the matrix specified with the data specified. In this case, cbind added a column to m with the numbers 101 to 105. Rbind would add another row to the matrix.

```{r import}
fly.worm <- read.delim("../Data/fly2worm.blastp.gz", header=F)
#change the file name and path to match whatever you used.
#the header=F argument tells R that the first row contains data, not column names
#takes a while to read in this large data set.
head(fly.worm)
summary(fly.worm)
colnames(fly.worm) <- c("qid", "sid", "pct", "len", "mis", "gaps", "qb", "qe", "sb", "se", "E", "S")
head(fly.worm)
summary(fly.worm)
```
##Exercise 8

```{r Exercise_8}
evalue <- sum(fly.worm["E"] == 0)
pID <- sum(fly.worm["pct"] > 50)
evalue_percent <- sum(fly.worm["E"] == 0)/nrow(fly.worm)
percentID <- sum(fly.worm["pct"] > 50)/nrow(fly.worm)
evalue_pID <- sum(fly.worm["E"] == 0 & fly.worm["pct"] < 50)
minpID <- subset(fly.worm, E == 0)
min(minpID["pct"])
```

**E-value of 0**: 7531 hits, 0.01494 
**Percent Identity > 50**: 20928 hits, 0.04151
**E-value of 0 and Percent Identity < 50**: 3351 hits
**Minimum Percent Identity of Hits with E-value of 0**: 24.07

##Exercise 9
I am surprised that low percent identity sequences can still have an E-value of 0. The alignment property that affects the E-value being 0 even when the percent identity is less than 50 is raw score. The higher the raw score, the lower the e-value. Thus, when the raw score is high enough, the e-value is small enough to get rounded to 0.

```{r Exercise_9}
mp <- cbind(fly.worm["E"], fly.worm["S"])
summary(mp)
mp_small <- subset(mp, E == 0)
mp_big <- subset(mp, E > 0)
mp_middle <- subset(mp, E < 2e-12)
mp_middle2 <- subset(mp, E < 2e-14)
summary(mp_small)
summary(mp_big)
summary(mp_middle)
summary(mp_middle2)
```
As the summaries show, the lower the E-values become, the higher the mean of the raw score "S" becomes. When E is 0 as shown in mp_small, the mean of the raw scores is many times higher than that if E is greater than 0 as seen in mp_big. Then as the high bounds for E decreeases, the mean of the raw score increases as seen in mp_middle and mp_middle2. 

##Exercise 10
```{r Exercise_10}
fly.worm.small <- sample(nrow(fly.worm), 10000)
summary(fly.worm[fly.worm.small, ])
```

##Exercise 11
```{r Exercise_11}
stereotype_db <- read.delim("stereotypes.csv", sep =",")
summary(stereotype_db)
head(stereotype_db)
```

##Activity Break for Intro to Plotting

###AB1-1
```{r AB1-1}
nerds_and_metal <- subset(stereotype_db, population == "nerd" | population == "metalhead")
summary(nerds_and_metal)
```

###AB1-2
```{r AB1-2}
males.only <- subset(stereotype_db, gender == "male" & beer > 25 & tacos > 20)
summary(males.only)
```
There are more males in the metalhead category who binge drink and eat over 20 tacos a week.

###AB1-3
```{r AB1-3}
print("How many females total spend more than 2 hours on the computer and drink more than 12 cups of coffee in a week?")
females.ask <- subset(stereotype_db, gender == "female" & computer > 2 & coffee > 12)
summary(females.ask)
```
161 females total.

###AB_4:
In R, "==" means if left hand side is equal to right side, then return true, else return false.

##Visualization with ggplot2 (Plotting tutorial)

###AB2-1

```{r AB2-1}
qplot(shower, data=stereotype_db, geom="histogram", fill=population, binwidth= 1)
```

###AB2-2
```{r AB2-2}
qplot(beer, data=stereotype_db, geom="histogram", fill=gender, binwidth= 1)
```

###AB2-3
```{r AB2-3}
ggplot(stereotype_db, aes(population, shower)) + geom_violin()
```

##Combining Subset with Visualization (Plotting tutorial)

###AB3-1
```{r AB3-1}
nerds_only <- subset(stereotype_db, population == "nerd")
qplot(beer, data=stereotype_db, geom ="density", fill=gender, alpha=I(0.5))
```

###AB3-2
```{r AB3-2}
qplot(coffee, data=stereotype_db, geom="freqpoly", colour = population, binwidth = 1)
metal <- subset(stereotype_db, population == "metalhead")
qplot(gender, tacos, data=metal, geom = "point", colour=tacos)
qplot(tacos,  data=stereotype_db, geom = "histogram", fill=population, binwidth=1)
```
Hipsters consume the most coffee per individual out of the entire population.
Of the metalheads, the males eat the most tacos and have a wider distribution.
Hippies eat the least amount of tacos.

###AB3-3
```{r AB3-3}
summary(stereotype_db)
qplot(coffee, data=stereotype_db, geom ="freqpoly", colour=population, binwidth=1)
qplot(tacos, data=stereotype_db, geom ="freqpoly", colour=population, binwidth=1)
qplot(shower, data=stereotype_db, geom="histogram", fill=population, binwidth=1)
qplot(computer, data=stereotype_db, geom="histogram", colour = population, binwidth=1)

```
Three hypotheses:
1. The more a person showers and goes on the computer, the more likely it is that they will be a nerd.
2. As the cups of coffee increase, the number of hipsters increase.
3. As the number of tacos increase, the amount of nerds, hippies, and hipsters decrease.

##Visualization with the ggplot function (Plotting tutorial)

###AB4-1
```{r AB4-1}
p <- ggplot(stereotype_db, aes(tacos, fill=gender))
p + geom_dotplot(binwidth=1) + facet_grid(~population, scales="free") 
```

###AB4-2
```{r AB4-2}
a <- ggplot(stereotype_db, aes(shower, colour=gender))
b <- ggplot(stereotype_db, aes(shower, fill=gender))
a + geom_bar()
b + geom_bar()
```

color= means to outline the edge of the unique attributes with different colors while fill= means to fill the whole bar or symbol with color, edges and all.

###AB5-1
```{r AB5-1}
c <- ggplot(stereotype_db, aes(beer, computer, color=population, shape=gender))
c+ geom_point()
```

###AB5-2
```{r AB5-2}
d <- ggplot(stereotype_db, aes(beer, fill=population, shape=gender), colour="clarity")
d + geom_bar() + scale_color_hue(l=90, c=50)
```

###AB5-3
```{r AB5-3}
f <- ggplot(stereotype_db, aes(gender, age))
f + geom_boxplot(aes(shape=population, fill=population), notch=TRUE, alpha=0.5)
```

##Visualization of the BLAST dataset

##Exercise 12
```{r Exercise_12}
fly.worm.sample <- sample(nrow(fly.worm), 10000)
fw.db <- fly.worm[fly.worm.sample, ]
summary(fw.db)
score.length <- ggplot(fw.db, aes(len, S))
score.length + geom_point(colour="red")
```

Using a random sampling of 10,000 points, we can roughly see the relationship between score and alignment length. The relationship is that as the alignment length increases, the score increases linearly.

##Exercise 13
Hypothesis: As the amount of gaps increase, the alignment length increases and as a result, the scores increase. So the bigger the gaps, the bigger the score.
```{r Exercise_13}
length.gaps <- ggplot(fw.db, aes(x=gaps, y=len))
length.gaps + geom_point(colour = "blue")
score.gaps <- ggplot(fw.db, aes(gaps, S))
score.gaps + geom_point(colour = "red")
```

Conclusion: As the amount of gaps increase, the alignment length does increase and thus the score increases. However, the increase in gaps does not directly result in the increase of score.

